Approximation as Differentiation

### 2025-06-09 main/approx-as-tangents.tex
comment about cartesian closure in L-poset

Smooth functions $f$ between manifolds $M$ and $N$ are functions on their points that are locally differentiable on $\RR^n$. Each smooth function induces maps of the (co)tangent spaces:
\begin{itemize}
\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$. In the Eucdlidean case when $M = \RR^m$ and $N = \RR^n$, the tangent map can be represented by the Jacobian matrix of partial derivatives of $f$ at $x$.
\item The \emph{backward derivative} (cotangent map, or pullback) $\pullf{f}_x$ is a linear map $\cotangents_{f(x)}(N) \linearto \cotangents_x(M)$. In the Euclidean case, the backward derivative is represented by the transpose of the Jacobian of $f$ at $x$.
\end{itemize}

@@

    In terms of \GPS the $\mathrm{gustave}$ function does not present a problem. For any particular run (i.e., input-output pair) of the program, there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.

@@

Even though stable functions can be defined on any partially ordered set, in light of the analogy with tangent spaces, it makes sense to require that meets (for forward approximation maps) and joins (for backwards approximation maps) exist:

@@

The following lemma is an instance of standard facts about Galois connections preserving meets and joins:

@@

@@

We now summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:

>>>>

Smooth functions $f$ between manifolds $M$ and $N$ are functions on their points that are locally differentiable on $\RR^n$. Manifolds and smooth functions form a category $\Man$. Each smooth function induces maps of the (co)tangent spaces:
\begin{itemize}
\item The \emph{forward derivative} (tangent map or pushforward) $\pushf{f}_x$ is a linear map $\tangents_x(M) \linearto \tangents_{f(x)}(N)$. In the Eucdlidean case when $M = \RR^m$ and $N = \RR^n$, the tangent map can be represented by the Jacobian matrix of partial derivatives of $f$ at $x$.
\item The \emph{backward derivative} (cotangent map, or pullback) $\pullf{f}_x$ is a linear map $\cotangents_{f(x)}(N) \linearto \cotangents_x(M)$. In the Euclidean case, the backward derivative is represented by the transpose of the Jacobian of $f$ at $x$.
\end{itemize}

@@

    In terms of \GPS the $\mathrm{gustave}$ function does not present a problem. For any particular run (i.e., input$\mapsto$output pair), there is an unambiguous minimal input that achieves the output, no matter that it was not achieved by a sequential processing of the input.

@@

Even though stable functions can be defined on any partially ordered set, in light of the analogy with tangent spaces, it makes sense to require that meets, preserved by forward approximation maps, and joins, preserved by backwards approximation maps, exist:

@@

This lemma is an instance of standard facts about Galois connections
preserving meets and joins:

@@

\bob{L-poset is has products and coproducts, but is not closed. To be closed would require each $\downset{x}$ to be a \emph{complete} lattice. However, this is tricky to implement as we need an infinitary operation to compute the reverse approximation map for function types. See \cite{amadio-curien} for the proof that bounded-complete lattices with stable functions form a cartesian closed category.}

@@

Let us summarise our analogy between differentiable functions between manifolds and stable functions between L-posets:
